---
title: "Creation of super regions"
author: "Martin Durocher"
date: "17/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
## Set colors for displaying the clusters
mycol12 <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c',
						 '#fdbf6f','#ff7f00','#cab2d6','#6a3d9a','#ffff99','#b15928')

col2pch <- function(z) z %/%12 + 16 
palette(mycol12)
```

## Introduction

When the record length of a site of interest is short, the uncertainty of estimated flood quantiles can be large.
In these situation, pooled frequency analysis may be used to reduce the uncertainty by transfering information from nearby sites to the target site.
Such pooling groups are formed according to a notion of distance, for floodNet recommend to use a measure of similarity based on to the seasonality of annual maximum discharges.
Canada has a vast diversity of flood regime and pool frequency analysis are commonly performed inside specific administrative regions. 
The idea of super regions was introduced to further enforce similarities that are not accounted for by the selected distance.
Indeed, limiting the pooling frequency analysis to specific administrative regions, such as provinces, can be seen as a form geographical super regions where sites considered in the analysis are geographically restricted.
The disavantage of such approach is that the administrative boundary are not generally obtained according to relevant hydrological properties.

The package `floodnetProject16` comes with a set of already selected gauged stations `gaugedSites` for which super regions based on are suggested. This document explained how they were obtained.

## Descriptors

We consider two spaces where the similarity between sites can be represented.
They is the geographical and the descriptor space.
The descriptor space has for coordinates the drainage area (AREA) and mean annual precipitation (MAP).
In Canada, AREA and MAP can be used as surrogate for climate and scale.
See Mostofi Zadeh and Burn (2019) for more details.
The graphic below shows the 1114 gauged stations of interest inside these two spaces.

```{r, fig.height=5, fig.width=10}

library(floodnetProject16)
library(CSHShydRology)

desc <- log(gaugedSites[,c('area','map')])
coord <- gaugedSites[,c('lon','lat')]

## Generic function to plot the geographical and physical spaces.
myfig <- function(col = 'black', pch = 16, legend.plot = TRUE){
  
  layout(matrix(c(1,2), 1,2))
  
  ucol <- sort(unique(col))
  
  plot(coord, pch = pch, col = col, ylim = c(42, 73),
       main = 'Geographical space', xlab = 'LON', ylab = 'LAT')
  
  if(legend.plot)
    legend('top', horiz = TRUE,legend = seq_along(ucol), col = ucol, pch = pch, cex = .8)
  
  plot(desc, pch = pch, col = col, 
  		 xlab = 'AREA (log)',
  		 ylab = 'MAP (log)',
  		 main = 'Descriptor space')
  
}

myfig(legend.plot = FALSE)
```

To create super regions, we consider the similarity based on the four descriptors: 
drainage area (AREA), mean annual precipitation (MAP), longitude (LON) and latitude (LAT).
We use log transformation on the AREA and MAP to obtain descriptors that are more concentrated around its central values.
We also use multidimensional scaling to project the geographical coordinates in a 2D space that will better preserve the great-circle when treated as Cartesian coordinates.

```{r}
## Project the coordinate using Multidimensional scaling
ecoord <- -cmdscale(GeoDist(coord),2)
colnames(ecoord) <- c('lon','lat')

xd <- cbind(scale(ecoord), scale(desc))
```

## Results

Hierarchical clustering techniques have often been used to delineate hydrologically homogenous regions.
Using more specifically the Ward's method, all sites are initially considered as their own cluster and are gradually paired up according to the similarity between clusters.
This creates a tree structure that is used to form meaningful groups of sites. 

The graphic below illustrates super regions defined by 6 and 12 clusters.
We see that the super regions are relatively compact in both the geographical space 
and the physical space.

```{r, fig.height=6, fig.width=12}
## Perform the classification

hc <- hclust(dist(xd), method = 'ward.D2')

## Function that reorgnanize the group label by size
OrderGrp <- function(z){
  lon <- tapply(coord[,1], z, median)
  mapid <- match(1:length(lon),order(lon))
  mapid[z]
}

## Identify 4 clusters
hc.grp6 <- OrderGrp(cutree(hc, k = 6))
hc.grp12 <- OrderGrp(cutree(hc, k = 12))

table(hc.grp6)
table(hc.grp12)

myfig(col = hc.grp6)
myfig(col = hc.grp12)

```

Another method that can be used for forming super regions is the k-mean clustering technique, where 
each cluster, or region, is defined by the set of the nearest stations. 
In the present context, a predifine number of region must be determined and a station is classified by finding the closet center. 
Initially, the algorithm start by selecting at random the centers and update them in order to minimize the intra-cluster variances.
The figures below illustrate super regions of size 6 and 12 using k-mean clustering.


```{r, fig.height=6, fig.width=12}
## Identify 4 clusters
set.seed(43)
km.grp6 <- OrderGrp(kmeans(xd, 6, nstart = 100)$cluster)
km.grp12 <- OrderGrp(kmeans(xd, 12, nstart = 100)$cluster)

myfig(col = km.grp6)
myfig(col = km.grp12)
```


```{r}
## Final set of 
supreg <- data.frame(
  supreg_hc6  = hc.grp6,
  supreg_hc12 = hc.grp12,
  supreg_km6  = km.grp6,
  supreg_km12 = km.grp12)

gaugedSites <- cbind(gaugedSites[,1:25], supreg)
```


## References

* Mostofi Zadeh, S., & Burn, D. H. (2019). A Super Region Approach to
Improve Pooled Flood Frequency Analysis. Canadian Water Resources Journal
/ Revue Canadienne Des Ressources Hydriques, 0(0), 1–14.
https://doi.org/10.1080/07011784.2018.1548946
